Concurrent Programming with CXL

### 2020-06-29 book/paper.tex
Implemented hash()

Mutual exclusion is an example of a \emph{safety property},
\index{safety property}%
a property that ensures that \emph{nothing bad will happen}, in this case
two processes being in the critical section.
What we need now a \emph{liveness property}:
\index{liveness property}%
we want to ensure that
\emph{eventually something good will happen}.
There are various possible liveness properties we could use,
but here we will propose the following informally: if there exists a non-empty
set $S$ of processes that are trying to enter the critical section and
processes in the intersection always leave eventually, then eventually one
process in $S$ will enter the critical section.
We call this \emph{progress}.
\index{progress}%

@@

\section*{Exercises}
\begin{problems}
\item Run Figure~\ref{fig:cs} using CXL.  As there is no protection of the critical
section, mutual exclusion is violated,
the assertion should fail, and a trace should be reported.
Now plug the following code \textbf{while} \texttt{True}: \textbf{pass};;
just before entering the critical section
in Figure~\ref{fig:cs} and run CXL again.
Mutual exclusion is guaranteed but progress is violated.
CXL should print a short trace
to a state from which a terminating state cannot be reached.
Describe in English the difference in the failure reports before
and after inserting the code.
\item See if you can come up with some different approaches that satisfy both
mutual exclusion and progress.  Try them with CXL and see if they work or not.
If they don't, try to understand why.
Do not despair if you can't figure it out---as
we will find out, it is possible but not easy.
\end{problems}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{$e$ == $e$}, \texttt{$e$ != $e$} & two CXL values of the same type \\
\hline
\texttt{$e$ < $e$, $e$ <= $e$} & any two CXL values \\
\hline
\texttt{$e$ > $e$, $e$ >= $e$} & any two CXL values \\
\hline
\texttt{$e$ and $e$ and ...} & two or more booleans \\
\hline
\texttt{$e$ or $e$ or ...} & two or more booleans \\
\hline
\texttt{not $e$} & a boolean \\
\hline
\texttt{-$e$} & an integer \\
\hline
\texttt{$e$ + $e$ + ...} & two or more integers, sets, or lists \\
\hline
\texttt{$e$ - $e$} & two integers or sets \\
\hline
\texttt{$e$ * $e$ * ...} & two or more integers or sets \\
\hline
\texttt{$e$ / $e$} & two integers (integer division) \\
\hline
\texttt{$e$ \% $e$} & two integers (division remainder) \\
\hline
\texttt{$e$~..~$e$} & two integers \\
\hline
\texttt{$e$ [not] in $e$} & first is any CXL value, second is set \\
\hline
\texttt{$e$ if $e$ else $e$} & middle $e$ must be a boolean \\
\hline
\texttt{\^{}$e$} & an address \\
\hline
\texttt{choose $e$} & a set \\
\hline
\texttt{min $e$, max $e$} & a set \\
\hline
\texttt{cardinality $e$} & a set \\
\hline
\texttt{keys $e$, len $e$} & a dictionary (includes tuple or list) \\
\hline
\texttt{bagsize $e$} & a bag \\
\hline
\texttt{atLabel $e$} & atom (corresponding to a label) \\
\hline
\texttt{nametag $e$} & $e$ must be \texttt{()} \\
\hline
\texttt{processes $e$} & $e$ must be \texttt{()} \\
\hline
\texttt{\&}\textit{lv} & \textit{lv} is an lvalue \\
\hline
\texttt{stop} \textit{lv} & \textit{lv} is an lvalue \\
\hline
\end{tabular}
\vspace{1em}

>>>>

Mutual exclusion is an example of a \emph{safety property},
\index{safety property}%
a property that ensures that \emph{nothing bad will happen}, in this case
two processes being in the critical section.
What we need now a \emph{liveness property}:
\index{liveness property}%
we want to ensure that
\emph{eventually something good will happen}.
There are various possible liveness properties we could use,
but here we will propose the following informally: if
(1) there exists a non-empty
set $S$ of processes that are trying to enter the critical section and
(2) processes in the critical section always leave eventually, then
eventually one process in $S$ will enter the critical section.
We call this \emph{progress}.
\index{progress}%

@@

\section*{Exercises}
\begin{problems}
\item Run Figure~\ref{fig:cs} using CXL.  As there is no protection of the critical
section, mutual exclusion is violated,
the assertion should fail, and a trace should be reported.
Now plug the following code \textbf{while} \texttt{True}: \textbf{pass};;
just before entering the critical section
in Figure~\ref{fig:cs} and run CXL again.
Mutual exclusion is guaranteed but progress is violated.
CXL should print a trace
to a state from which a terminating state cannot be reached.
Describe in English the difference in the failure reports before
and after inserting the code.
\item See if you can come up with some different approaches that satisfy both
mutual exclusion and progress.  Try them with CXL and see if they work or not.
If they don't, try to understand why.
Do not despair if you can't figure it out---as
we will find out, it is possible but not easy.
\end{problems}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{$e$ == $e$}, \texttt{$e$ != $e$} & two CXL values of the same type \\
\hline
\texttt{$e$ < $e$, $e$ <= $e$} & any two CXL values \\
\hline
\texttt{$e$ > $e$, $e$ >= $e$} & any two CXL values \\
\hline
\texttt{$e$ and $e$ and ...} & two or more booleans \\
\hline
\texttt{$e$ or $e$ or ...} & two or more booleans \\
\hline
\texttt{not $e$} & a boolean \\
\hline
\texttt{-$e$} & an integer \\
\hline
\texttt{$e$ + $e$ + ...} & two or more integers, sets, or lists \\
\hline
\texttt{$e$ - $e$} & two integers or sets \\
\hline
\texttt{$e$ * $e$ * ...} & two or more integers or sets \\
\hline
\texttt{$e$ / $e$} & two integers (integer division) \\
\hline
\texttt{$e$ \% $e$} & two integers (division remainder) \\
\hline
\texttt{$e$~..~$e$} & two integers \\
\hline
\texttt{$e$ [not] in $e$} & first is any CXL value, second is set \\
\hline
\texttt{$e$ if $e$ else $e$} & middle $e$ must be a boolean \\
\hline
\texttt{\^{}$e$} & an address \\
\hline
\texttt{hash $e$} & any CXL value \\
\hline
\texttt{choose $e$} & a set \\
\hline
\texttt{min $e$, max $e$} & a set \\
\hline
\texttt{cardinality $e$} & a set \\
\hline
\texttt{keys $e$, len $e$} & a dictionary (includes tuple or list) \\
\hline
\texttt{bagsize $e$} & a bag \\
\hline
\texttt{atLabel $e$} & atom (corresponding to a label) \\
\hline
\texttt{nametag $e$} & $e$ must be \texttt{()} \\
\hline
\texttt{processes $e$} & $e$ must be \texttt{()} \\
\hline
\texttt{\&}\textit{lv} & \textit{lv} is an lvalue \\
\hline
\texttt{stop} \textit{lv} & \textit{lv} is an lvalue \\
\hline
\end{tabular}
\vspace{1em}
